TacticToe: Learning the Game of Tactics

### 2018-01-12 tactictoe_jv.tex
reformulations in the abstract

\begin{abstract}
Formalization of mathematical statements in interactive theorem provers (ITPs) 
relies nowadays on many methods to automate the construction of a proof.
Some of this method are general purpose and with enough time can prove any 
theorems. Yet, in a formalized proof script 
a mathematician take advantages of a combination of specialized rules, 
theory-based strategy and general ones. All of them are implemented as tactics 
in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
strategies. Sometimes, a strategy can be specialized for only one theory as in 
SMT solving. But it is difficult for an automated theorem prover to rely on 
rules tailored for a particular case. 
Indeed, this would require implementing many specific rules and more 
importantly figuring out in which cases they can be applied.

We bridge this gap by implementing our tactical prover TacticToe on top 
of the ITP HOL4. 
After learning in which situations tactics are effective from 
tactic-level proof and what kind of formulas are the results of effective 
steps, we implement a Monte Carlo tree search algorithm that explores the most 
promising tactic-level proof paths. Experiments on the \holfour standard 
library shows that TacticToe is able to re-prove 59\% of 7088 theorems in less 
than 10 seconds where as the ATP Eprover with auto-schedule
solves 29\%.

>>>>

\todo{JU: Still feels more like an intro than an abstract. Could be shorter and punchier.}
\begin{abstract}
Formalization of mathematics in interactive theorem provers (ITPs) 
relies today on many proof automation methods.
Some of these methods are general-purpose and with enough time they can prove any 
theorem. Many formalized proof scripts however also
take advantage of specialized rules and theory-based strategies. 
In the HOL4 ITP system, practically all of them are implemented as \emph{tactics}.

Automated theorem provers (ATPs) rely on parameterizable general 
strategies. Sometimes, a strategy can be specialized for only one theory as in 
SMT solving. But it is difficult for an automated theorem prover to rely on 
rules tailored for a particular case. 
Indeed, this would require implementing many specific rules and more 
importantly figuring out in which cases they can be applied.

We bridge this gap by implementing our tactical prover TacticToe on top 
of the HOL4 ITP. 
TacticToe first learns from many proof scripts
the relevance of tactics in particular proof situations.
% are effective from 
% tactic-level proofs and what kind of formulas are the results of effective 
% steps, 
This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths. 
Experiments on the \holfour standard 
library show that TacticToe is able to re-prove 59\% of the 7088 theorems in less 
than 10 seconds compared to 29\% proved by the HolyHammer system using the E ATP in its auto-schedule mode.
